---
title: "Introduction to LimpiaR"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
   
vignette: >
  %\VignetteIndexEntry{Introduction to LimpiaR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# LimpiaR Overview

LimpiaR is a package built to expedite pre-processing and cleaning of text data, with some handy functions in Spanish and R. To get started, we'll load few helpful libraries.

# Walkthrough

```{r setup, echo = TRUE, results = "hide", warning = FALSE, message = FALSE}
library(magrittr)
library(dplyr)
library(stringr)
library(LimpiaR)
```

LimpiaR's functions begin with limpiar_. Once the library has been loaded, typing limpiar_ in an Rstudio script or Rmarkdown code block, will produce a drop down menu of all LimpiaR functions, which should help you to find the name of the function you're looking for - you can then use tab to autocomplete the function. Once inside the function, RStudio should give you a popover which shows the argument the function expects. You can also type 'control + space' when your cursor is inside the function's brackets to force extra help.

```{r, eval = TRUE, echo = FALSE, warning= FALSE}
data <- data.frame(
  `Mention Url` = c("www.twitter.com/post1",
                  "www.twitter.com/post2",
                  "www.facebook.com/post1",
                  "www.facebook.com/post2",
                  "www.twitter.com/post3",
                  "www.youtube.com/post1",
                  "www.instagram.com/post1",
                  "www.youtube.com/post2",
                  "www.youtube.com/post3",
                  "www.youtube.com/post4"),
  `Mention Content` = c(  "holaaaaaa! c贸贸mo    est谩s @magdalena   ?!",
                      "  han visto este articulo!? Que horror! https://guardian.com/emojisbanned NO SE PUEDE!!",
                      "ayyyyyy a mi me   gustaria ir a londres yaaa #llevame #porfavor",
                      "jajajajaja eres un wn!",
                      "RT dale un click a ver una mujer baila con su perro",
                      "grax ntonces q?",
                      "yo soy el mejor , no eres nada!! ぃぃ ",
                      "grax ntonces q?",
                      "grax ntonces q?",
                      "grax ntonces q?"))%>%
  tibble()%>%
  relocate(Mention.Content)
```

```{r}
data
```

### Column Names

We created a data frame of posts and URLs. After loading libraries and data, the first part of any workflow should be to clean the column names, this makes using tab completion, and accessing column names much faster (which in the long run = big productivity gains). For this, we'll use the janitor package. You can uncomment the code to install janitor if it is not already installed on your machine.

```{r}
# ifelse(!"janitor" %in% installed.packages(),
#    install.packages("janitor"), library(janitor))

(data <- data %>% 
   janitor::clean_names())

```

## Lower Case Text Variable

For most workflows, the next step will be to make the text variable lower case. We do this to make tokens like 'AMAZING' or 'Amazing' -> 'amazing'. You do not need a LimpiaR function for this, as the base R function tolower() works just fine.
```{r}
(data <- data %>%
  mutate(mention_content = tolower(mention_content)))
```


## Limpiar Functions

## limpiar_accents

Now we're going to look at LimpiaR's functions individually. The first function is limpiar_accents, this will replace the accents most common in Spanish words from the text variable, with their Latin-alphabet equivalents e.g. '茅 -> e' . We will use the assignment operator to make sure these changes are saved. 

Tip: you can type ?limpiar_accents to access the documentation, and see which arguments you need to fill in.

```{r}
(data <- data %>%
  limpiar_accents(text_var = mention_content))
```

### limpiar_duplicates

Now we'll remove duplicate posts, notice that we don't actually need to type text_var = mention_content, because the default argument for text_var is already mention_content. 

```{r}
(data <- data %>%
  limpiar_duplicates())
```

>Note: If the text column in our data frame were called 'text' we would have to specify text_var = text, or call:

```{r}
data %>% rename(text = mention_content)
```

### limpiar_retweets
If you need to remove retweets, for example to create a bigram network, limpiar has a function just for that. 
```{r}
(data <- data %>% 
   limpiar_retweets())
```

### limpiar_url
We generally don't want URLs appearing in our charts or analyses, so we can remove them with the limpiar_url function.

```{r}
(data <- data %>%
   limpiar_url())
```


### limpiar_spaces
Next we'll look at how to use LimpiaR to remove annoying white spaces, like those at the beginning of a sentence, or between punctuation, or multiple white spaces for no reason; as is common in the messy data we often encounter.

```{r}
(data <- data %>%
  limpiar_spaces())
```

### limpiar_tags
We can also remove user handles (e.g. @magdalena) and hashtags with the limpiar_tags function. Remember, you can type ?limpiar_tags to access documentation.

Replace only hashtags:
```{r}
data %>%
  limpiar_tags(user = FALSE, hashtag = TRUE)
```

Replace only user tags:
```{r}
data %>%
  limpiar_tags(user = TRUE, hashtag = FALSE)

```

Replace both hashtags and user handles:
```{r}
data %>%
  limpiar_tags()
```


### Quick recap - we've looked at:

* cleaning column names with janitor::clean_names()
* making the text variable lower case with mutate() &  tolower() 
* cleaning accents with limpiar_accents()
* cleaning duplicate posts with limpiar_duplicates()
* cleaning retweets with limpiar_retweets()
* cleaning urls with limpiar_url()
* cleaning spaces with limpiar_spaces()
* cleaning user handles and hashtags with limpiar_tags()

### limpiar_shorthands
One of the biggest problems with the messy data we encounter, are shorthands. Generally, algorithms have been trained on clean, standard language, so they do not encounter shorthands and abbreviations. Shorthands also change all the time, making it impractical to continuously train algorithms as new shorthands arise. This function attempts to bridge that gap, by normalising the most common shorthands.

```{r}
(data <- data %>%
   limpiar_shorthands())
```

### limpiar_repeated_chars
We don't want our algorithm to have to learn the difference between 'ajajaj' and 'jaja' or 'ay' and 'ayyyy' as practically speaking, there is none. We also don't want to introduce unnecessary tokens, so we normalise the most common occurrences of repeated or additional characters.

```{r}
(data <- data %>%
   limpiar_repeat_chars())
```

Generally, the steps we've taken so far will be used in each and every analysis/project to help clean the data. We will now look at some of the more circumstantial functions, i.e. they will not be used in every analysis.

### Emojis

Emojis are a type of non-ASCII unicode character, which means that removing all non-ASCII characters will remove emojis by force. It also means that functions designed to target certain unicode characters by patterns, may inadvertently remove other special characters - as well as emojis, or instead of emojis!

#### limpiar_recode_emojis()

<details>
<summary>Why not juse a regular expression? </summary>
We scraped some lists of emojis which get the emojis directly for replacement. This approach is more computationally expensive than filtering out via RegEx, but it is often more precise as it targets the emojis directly.
</details>

We don't need to use limpiar_recode_emojis for every analysis, as many ParseR & SegmentR functions ignore them implicitly. However, if we know that we need to replace them with their text descriptios, we can can use limpiar__recode_emojis(). One problem with this, and the reason it is for special cases only, is the emoji's encodings are in English. We may, at some point, translate them to Spanish, but it seems unlikely.

```{r}
data %>%
  limpiar_recode_emojis(text_var = mention_content, with_emoji_tag = FALSE)
```

Or if we set `with_emoji_tag` to TRUE, our emojis are now pasted together with '_' and have an '_emoji' label.
```{r}
data %>%
  limpiar_recode_emojis(mention_content, with_emoji_tag = TRUE)
```

> **Warning** the `limpiar_recode_emojis()` function is quite slow and scales poorly with the size of inputs. So if using on a large dataset with many long documents expect functions to take a while to run.

#### limpiar_remove_emojis()

What about in situations where we don't want to replace emojis with their text inputs, **or** we don't mind risking the loss of some other non-ASCII characters, **or** we want something that runs fast?

Instead of `limpiar_recode_emojis` we can use `limpiar_remove_emojis`! This function operates with a fairly simple RegEx pattern, meaning it runs a lot more efficiently than its recode counterpart.

```{r}
data %>%
  limpiar_remove_emojis(mention_content)
```

### non-ASCII characters

ASCII (American Standard Code for Information Interchange)  is a character-encoding standard for representing numbers and text. There are 128 ASCII characters,  including the letters from a-z in upper and lowercase, numbers 0-9, common punctuation marks, and a few additional characters with specific uses for computers. Everything else is non-ASCII.

<details>
<summary>All 128 ASCII characters</summary>
Dec  Hex  Char  Description
--- ----- ----- -----------
000  00   NUL   Null
001  01   SOH   Start of Heading
002  02   STX   Start of Text
003  03   ETX   End of Text
004  04   EOT   End of Transmission
005  05   ENQ   Enquiry
006  06   ACK   Acknowledge
007  07   BEL   Bell
008  08   BS    Backspace
009  09   HT    Horizontal Tab
010  0A   LF    Line Feed
011  0B   VT    Vertical Tab
012  0C   FF    Form Feed
013  0D   CR    Carriage Return
014  0E   SO    Shift Out
015  0F   SI    Shift In
016  10   DLE   Data Link Escape
017  11   DC1   Device Control 1 (XON)
018  12   DC2   Device Control 2
019  13   DC3   Device Control 3 (XOFF)
020  14   DC4   Device Control 4
021  15   NAK   Negative Acknowledge
022  16   SYN   Synchronous Idle
023  17   ETB   End of Transmission Block
024  18   CAN   Cancel
025  19   EM    End of Medium
026  1A   SUB   Substitute
027  1B   ESC   Escape
028  1C   FS    File Separator
029  1D   GS    Group Separator
030  1E   RS    Record Separator
031  1F   US    Unit Separator
032  20   SPACE Space
033  21   !     Exclamation Mark
034  22   "     Double Quote
035  23   #     Number Sign
036  24   $     Dollar Sign
037  25   %     Percent
038  26   &     Ampersand
039  27   '     Single Quote
040  28   (     Left Parenthesis
041  29   )     Right Parenthesis
042  2A   *     Asterisk
043  2B   +     Plus
044  2C   ,     Comma
045  2D   -     Hyphen
046  2E   .     Period
047  2F   /     Forward Slash
048  30   0     Zero
049  31   1     One
050  32   2     Two
051  33   3     Three
052  34   4     Four
053  35   5     Five
054  36   6     Six
055  37   7     Seven
056  38   8     Eight
057  39   9     Nine
058  3A   :     Colon
059  3B   ;     Semicolon
060  3C   <     Less Than
061  3D   =     Equals
062  3E   >     Greater Than
063  3F   ?     Question Mark
064  40   @     At Sign
065  41   A     Uppercase A
066  42   B     Uppercase B
067  43   C     Uppercase C
068  44   D     Uppercase D
069  45   E     Uppercase E
070  46   F     Uppercase F
071  47   G     Uppercase G
072  48   H     Uppercase H
073  49   I     Uppercase I
074  4A   J     Uppercase J
075  4B   K     Uppercase K
076  4C   L     Uppercase L
077  4D   M     Uppercase M
078  4E   N     Uppercase N
079  4F   O     Uppercase O
080  50   P     Uppercase P
081  51   Q     Uppercase Q
082  52   R     Uppercase R
083  53   S     Uppercase S
084  54   T     Uppercase T
085  55   U     Uppercase U
086  56   V     Uppercase V
087  57   W     Uppercase W
088  58   X     Uppercase X
089  59   Y     Uppercase Y
090  5A   Z     Uppercase Z
091  5B   [     Left Bracket
092  5C   \     Backslash
093  5D   ]     Right Bracket
094  5E   ^     Caret
095  5F   _     Underscore
096  60   `     Backtick
097  61   a     Lowercase a
098  62   b     Lowercase b
099  63   c     Lowercase c
100  64   d     Lowercase d
101  65   e     Lowercase e
102  66   f     Lowercase f
103  67   g     Lowercase g
104  68   h     Lowercase h
105  69   i     Lowercase i
106  6A   j     Lowercase j
107  6B   k     Lowercase k
108  6C   l     Lowercase l
109  6D   m     Lowercase m
110  6E   n     Lowercase n
111  6F   o     Lowercase o
112  70   p     Lowercase p
113  71   q     Lowercase q
114  72   r     Lowercase r
115  73   s     Lowercase s
116  74   t     Lowercase t
117  75   u     Lowercase u
118  76   v     Lowercase v
119  77   w     Lowercase w
120  78   x     Lowercase x
121  79   y     Lowercase y
122  7A   z     Lowercase z
123  7B   {     Left Brace
124  7C   |     Vertical Bar
125  7D   }     Right Brace
126  7E   ~     Tilde
127  7F   DEL   Delete

</details>

For our purposes we have extended the ASCII characters to include Latin accents (茅, 铆, etc.), let's get our original data frame back to demonstrate. We should remove things like emojis, but keep our punctuation and accented characters:

```{r, eval = TRUE, echo = FALSE, warning= FALSE}
data <- data.frame(
  `Mention Url` = c("www.twitter.com/post1",
                  "www.twitter.com/post2",
                  "www.facebook.com/post1",
                  "www.facebook.com/post2",
                  "www.twitter.com/post3",
                  "www.youtube.com/post1",
                  "www.instagram.com/post1",
                  "www.youtube.com/post2",
                  "www.youtube.com/post3",
                  "www.youtube.com/post4"),
  `Mention Content` = c(  "holaaaaaa! c贸贸mo    est谩s @magdalena   ?!",
                      "  han visto este articulo!? Que horror! https://guardian.com/emojisbanned NO SE PUEDE!!",
                      "ayyyyyy a mi me   gustaria ir a londres yaaa #llevame #porfavor",
                      "jajajajaja eres un wn!",
                      "RT dale un click a ver una mujer baila con su perro",
                      "grax ntonces q?",
                      "yo soy el mejor , no eres nada!! ぃぃ ",
                      "grax ntonces q?",
                      "grax ntonces q?",
                      "grax ntonces q?"))%>%
  tibble()%>%
  relocate(Mention.Content) %>%
  janitor::clean_names()

data
```

```{r}
data %>%
  limpiar_non_ascii(mention_content)
```


### limpiar_alphanumeric
Similar to non-ASCII characters, we can retain only the alphanumeric characters (a-zA-Z0-9 + spaces). This is a heavy-duty option which *will* remove all accented characters.
```{r}
data %>%
  limpiar_alphanumeric(mention_content)
```

If we want to use `limpiar_alphanumeric` **and retain** our accented characters, then we should recode the accents first with `limpiar_accents`:
```{r}
data %>%
  limpiar_accents(mention_content) %>%
  limpiar_alphanumeric(mention_content)
```

> You'll need to make an informed choice between `limpiar_alphanumeric`, `limpiar_non_ascii` and other functions like `limpiar_accents` and `limpiar_*_emojis`

### limpiar_stopwords
Stop words are common words that do not provide us with much information as to an utterance's meaning. For example, in the sentence: 'the man is in prison for theft', if we knew only one word from this sentence, and that word was 'is', 'in', 'the', or 'for' then we wouldn't have much idea what the sentence is about. However, 'prison' or 'theft', would give us a lot more information.

For many analyses, we remove stop words to help us see the 'highest information' words, to get a high-level understanding of large bodies of texts (such as in topic modelling and bigram networks.) For virtually all scenarios, you will want to use the limpiar_stopwords() with the argument stop_words = "topics" like so:

```{r}
data %>%
  limpiar_stopwords(stop_words = "topics") %>%
  limpiar_spaces() #to clear the spaces of words that were removed
```

However, sometimes we want to keep words that would usually be treated as stopwords for a specific purpose. For example, when we're analysing sentiment 'negatives' can invert the sentiment of a text - 'no me gusta' vs 'me gusta'. If we remove all instances of 'no' from our data, we will do a worse job at analysing sentiment. For Spanish we have a slightly shorter list of stopwords for sentiment than topics, where we have removed a few choice terms.
```{r}
data %>%
  limpiar_stopwords(stop_words = "sentiment") %>%
  limpiar_spaces() 
```

Warning - sentences can look quite strange without stopwords, and a lot of social posts are virtually meaningless altogether! 

> It's also worth pointing out, that a lot of information can be lost when removing stop words. Many phrases in English and Spanish have very different meanings when a stop word is removed, and some stopwords lists contain negatives, which can drastically change the meaning of a sentence! So use with care.

## Utility Functions
We are nearly at the end of this introduction to LimpiaR, but before we finish, let's look at two utility functions which may be useful. We've conjoured up a new data frame called df, which we will use to show the last two functions and how to chain everything together.

```{r, eval = TRUE, echo = FALSE, warning= FALSE}
df <- data.frame(
  `Mention Url` = c("www.twitter.com/post1",
                  "www.twitter.com/post2",
                  "www.facebook.com/post1",
                  "www.facebook.com/post2",
                  "www.twitter.com/post3",
                  "www.youtube.com/post1",
                  "www.instagram.com/post1",
                  "www.youtube.com/post2",
                  "www.youtube.com/post3",
                  "www.youtube.com/post4"),
  `Mention Content` = c(  "holaaaaaa! c贸贸mo    est谩s @magdalena   ?!",
                      "  han visto este articulo!? Que horror! https://guardian.com/emojisbanned NO SE PUEDE!!",
                      "ayyyyyy a mi me   gustaria ir a londres yaaa #llevame #porfavor",
                      "jajajajaja eres un wn!",
                      "RT dale un click a ver una mujer baila con su perro",
                      "grax ntonces q?",
                      "yo soy el mejor , no eres nada!! ぃぃ ",
                      "grax ntonces q?",
                      "grax ntonces q?",
                      "grax ntonces q?"),
  na_col = c(NA, NA, NA, NA, NA, NA, NA, NA, "tadaa", NA))%>%
  tibble()%>%
  relocate(Mention.Content)%>%
  janitor::clean_names()
```
 
```{r}
df
```
### limpiar_inpsect
So, imagine that we see a strange pattern, and we want to check what's going on with that specific pattern. We can use limpiar_inspect to view all posts which contain that pattern in an interactive frame!

```{r, eval = FALSE}
limpiar_inspect(df, 
                pattern = "ntonces", 
                text_var = mention_content,
                url_var = mention_url,
                title = "ntonces")
```


```{r, eval = TRUE, echo = FALSE}
limpiar_inspect(df, 
                pattern = "ntonces", 
                text_var = mention_content,
                url_var = mention_url, 
                title = "ntonces") %>%
  knitr::kable() 
```

Whilst it's pretty obvious that all of the 'grax ntonces q?' posts are exactly the same, in the real world we're going to have 10,000 times as many posts, and searching for suspicious patterns may take up a lot of our time. 

### limpiar_na_cols
This final function is useful when we want to remove 'mostly NA' columns of a data frame. We may want to do this to save memory, for example if we have 400,000 posts and 80 columns. In this case we'll get rid of all columns for which 25% or more of their values are NA.

```{r}
limpiar_na_cols(df,threshold =  0.25)
```

## Putting It All Together

To speed things up, we **could** call the functions together in one big long pipe.  
```{r}
df %>%
  limpiar_na_cols(threshold = 0.25)%>%
  limpiar_accents()%>%
  limpiar_retweets()%>%
  limpiar_shorthands()%>%
  limpiar_repeat_chars()%>%
  limpiar_url()%>%
  limpiar_remove_emojis()%>%
  limpiar_shorthands()%>%
  limpiar_spaces()%>%
  limpiar_duplicates()
```

However, we generally want to check what the effects of our transformations are on our data, so doing a lot of operations like this without any intermediate checks is risky. 













